Measuring Software Ticket Quality using Quantitative Data Analysis

### 2018-04-22 paper/paper.tex
Fix issues and typos

We split our tool into four main commands: \emph{store}, \emph{analyse}, \emph{plot} and \emph{stats}. They all 
follow the flow shown in figure \ref{fig-eg} and complete the whole application cycle, in the end producing a database 
of analysed tickets, as well as plots and statistics for investigating correlations, saved on the filesystem.

@@

The goroutines interact with the Jira Server REST API. The goroutines exploit the pagination facility in the API, which means 
slicing the whole set of tickets into multiple smaller arrays to improve performance and reduce the load on the server. We also
specified to Jira exactly what set of issue fields to return and, while responses were retrieved from the instance,
we began storing them into our Bolt DB instance.

The way we chose to store them was to set the issue key (i.e. unique identifier that differentiates a ticket from all others across all 
projects inside a Jira instance) as the key of the pair and the value is the JSON representation of the ticket. This storage method, in conjunction
with how simple Bolt is designed, helped us reduce the size of the database significantly. For small databases, 
using RDBMSs (e.g. Postgres, MySQL, Oracle) is not performance critical, but when the data set is large such as in our study, bottlenecks might start affecting 
the progress of the project. This is mainly due to the structure of how Jira returns the tickets; for example, all issues have comments as 
a field, consisting of an array of structs with body, author, created time etc. Then, each comment has a changelog history attached to it, 
having in turn multiple changelog groups and also each group can have multiple changelog items. Therefore, this long pipeline became a serious 
bottleneck for us once we reached a certain number of tickets because we had relations between all these tables, compared to Bolt which does not 
link any key-value pair to another. Another major point when we decided to proceed with Bolt was that we did not need to query the data at all 
as all our processing was too complex to convert the logic into an SQL query.

@@

Once the Bolt DB instance is completely populated with all the tickets we were interested in, the analyze command first 
fetches all the issues in the database. The next step is to filter out the tickets based on the following criteria:
\begin{itemize}
  \item tickets are \emph{High Priority} - the ticket status was marked either Blocker, Critical, Major or High;
  \item tickets were closed and had a maximum \emph{Time-To-Close} value of 27,000 hours (roughly 3 years) (we added the 
  higher bound as well in order to eliminate outliers retrieved by Jira);
  \item remove outliers for all categories - some fields for a small number of tickets that had extreme values (e.g. one ticket
  had over 270,000 comments) introduced by the team behind the project and we needed to exclude them so that the analysis 
  produced correct results.
\end{itemize}

Then, the analysis routine runs in parallel the \emph{seven types of analysis} for answering the research questions 
outlied in Section \ref{intro}, each in its own separate goroutine in order to speed up the process:
\begin{itemize}
  \item \emph{attachments} - checks whether attachments influence Time-To-Close and also look at what types 
  of attachments (e.g. screenshots, archive, code snippets) influence time-to-close the most;
  \item \emph{steps to reproduce} - performs a complex regex to detect steps to reproduce and then checks whether their presence
  influence the time-to-close of a ticket; the regular expression was adapted after the general guidelines elicited by 
  Bettenburg et al. \cite{bettenburg2008extracting} on how to extract structured data, including enumerations (of which 
  steps to reproduce are a subset);
  \item \emph{stack traces} - runs complex regex for detecting exception stack traces in either summary, description or comments 
  and verifies whether their presence influence Time-To-Close for the ticket; this analysis only inspects projects
  written in Java as it is the most popular proponent built-in of stack traces;
  \item \emph{comments complexity} - loops through all comments and counts all words; then, the tool verifies whether 
  the number of words in comments has an impact on the Time-To-Close for the ticket;
  \item \emph{fields complexity} - same analysis as comments complexity, but only for summary and description wordiness;
  \item \emph{grammar correctness} - it uses Azure Cloud Bing Spell Check API to perform analysis on summary, 
  description and comments; after concatenating everything and making it compatible with the API allowed formats, the tool 
  receives back in the JSON payload not only the number of flagged tokens (i.e. grammar errors), but also their types 
  (e.g. unknown token, misspell);
  \item \emph{sentiment} - uses Google Cloud Platform's Natural Language Processing API to retrieve the 
  sentiment score for summary, description and comments; it first concatenates everything and makes it conform to Google's 
  API and then sends the request, receiving a score from -1 to 1 inclusive (-1 means most negative, 1 means most 
  positive);
\end{itemize}

@@

In this section we present our results to the research questions we set to answer. We have computed charts for all 
independent variables using either scatter plots for continuous variables (e.g. number of comments/time-to-close) 
or bar charts for categorical variables (e.g. presence of attachments). We also computed statistical analysis for 
all variables to test their significance and rank correlation coefficient and we will present each of them in their 
corresponding subsection.

The answers to our research questions are all statistically significant and they can summarised as follows:
\begin{itemize}
  \item having attachments reduces time-to-close of a ticket;
  \item having steps to reproduce reduces time-to-close of a ticket;
  \item having stack traces reduces time-to-close of a ticket;
  \item more grammar errors imply increase time-to-close values;
  \item lower sentiment scores result in increased time-to-close values;
  \item verbose comments result in increased time-to-close values;
  \item verbose summary and description imply increased time-to-close values.
\end{itemize}

@@

Using the plot package in our Go tool, we automatically generated the bar chart shown in Figure \ref{attachments}. 
As it can be seen, various types of attachments can produce completely different \emph{times-to-close}. For 
tickets with archive we can see the longest time-to-close mean while tickets having text attachments are at 
the other end of the spectrum. The smallest mean is for other attachments, which is represented by any file 
that does not have an extension specified in one of the other categories.

We assume that image attachments are screenshots attached to the ticket in order to help developers see and 
subsequently reproduce the bug. Thus, we can clearly see that the difference between the means of time-to-close 
for tickets without attachments and tickets with screenshots is significantly in favor of the latter. This can 
indicate that developers indeed find screenshots helpful in solving the ticket faster.

Another insightful result is the fact that having text, config and spreadsheet attachments significantly 
reduces the time-to-close for the ticket. This can imply that developers find value in various steps to reproduce, 
config values used by users when the bug encountered or data in a spreadsheet format (e.g. csv) that can be 
found in such files and helps them lower the time-to-close for the ticket.

On the other hand, having no attachments does not necessarily mean that the ticket will get closed after 
a longer period of time. As it can be seen in the figure, the mean time-to-close value for tickets without 
attachments is somewhat neutral in the entire data set.

@@

Thus, our analysis shows that having attachments in most formats, ranging from markdown files to PNG screenshots 
and snippets of code, can help developers solve the tickets faster. We believe that the solution of making attachments
provide the most value to the software project would be if issue tracking systems would make explicit to the reporters 
that uploading a screenshot or a code snippet can help reduce the time-to-close for the ticket.

@@

As it can be shown in the figure, having steps to reproduce can significantly reduce the time-to-close 
for the ticket. Even though the data collected does not have such many tickets with steps to reproduce 
(39,988 compared to the total number of 238,286 closed tickets), it is still a valuable finding which shows 
that steps to reproduce are indeed a quality factor for software tickets.

@@

Thus, the answer to our research question is that the presence of steps to reproduce in tickets 
has a positive influence on the time-to-close. One solution that we propose for maximizing their benefit 
is that issue tracking systems should include hints on the ticket creation form specifying that 
steps to reproduce should be included as developers would be able to fix the bug faster. Moreover, 
issue tracking systems could also automatically add a new system field (like summary or description) 
for any project created and the project maintainers or administrators could make them mandatory for 
when a ticket is created by reporters.

@@

As shown in Figure \ref{stack_traces}, we can see that the presence of exception stack traces can 
reduce the time-to-close of a ticket in orders of magnitude. This could imply that stack traces help 
developers working on the project localise the code that is running exceptions

@@

A correlational analysis ran on all 201,786 tickets, however, releaved a weak positive relationship with $r = 0.271$ and $p < 0.01$. 
This implies a statistically significant finding - as the number of comments increases, the time-to-close value of the ticket also increases.

@@

Through calling Microsoft Azure's Bing Spell Check API, one of the best grammar analysis APIs in the industry, we 
managed to calculate the total number of grammar errors in summary, description and comments for a total number 
of 133,689 tickets. As shown in a number of studies, the number of spelling errors can influence the quality of a ticket. 
However, in the work conducted by Schugerl et al. \cite{schugerl2008mining}, the authors found that spelling errors 
can be found in both high and low quality tickets, thus should not be considered as a reliable factor for determining
the Time-To-Close for a ticket. In the rest of this section we will demonstrate the contrary - tickets with more 
grammar errors (spelling errors are just a subset of the types of errors we are investigating) imply lower quality
tickets (i.e. increased time-to-close).

@@

On the other side of the spectrum, having a more positive score leads to reduced time-to-close values for tickets. Calefato et al. 
\cite{calefato2015mining} conducted a study where they show that a positive sentiment has the biggest impact on 
the quality of an answer on Stack Overflow, one of the most popular websites related to programming in general. This 
implies that when developers and triagers treat their reporters respectfully and offer support whenever possible, 
the users will feel more welcomed and willing to help report other bugs in the future as well. We also have scores of 
over 0.9, meaning that the actors participating in the discussion (or the reported summary/description) might compliment 
each other for doing a great job.

@@

We manager to answer all seven research questions and find factors that can posittively or 
negatively affect the time-to-close value of a ticket. All of our results are statistically significant and we can summarise 
them as follows:
\begin{itemize}
  \item the presence of steps to reproduce reduces the time-to-close for a ticket;
  \item the presence of stack traces reduces the time-to-close for a ticket;
  \item the presence of attachments reduces the time-to-close for a ticket - more specifically tickets with screenshots, 
  config files or text files help tickets get closed quicker than those without attachments;
  \item higher numbers of grammar errors imply increased time-to-close values;
  \item higher sentiment scores imply reduced time-to-close values for a ticket;
  \item higher numbers of words in comments imply increased time-to-close values for a ticket;
  \item higher number of words in description and summary increase the time-to-close values for tickets.
\end{itemize}

>>>>

We split our tool into four main commands: \emph{store}, \emph{analyse}, \emph{plot} and \emph{stats}. By following the flow shown in 
Figure \ref{fig-eg} and completing the whole application cycle, a database of analysed tickets is produced,
as well as plots and statistics for investigating correlations, saved on the filesystem.

@@

The goroutines interact with the Jira Server REST API. The goroutines exploit the pagination facility in the API, which means 
slicing the whole set of tickets into multiple smaller arrays to improve performance and reduce the load on the server. We also
specified to Jira exactly what set of issue fields to return and, as responses were retrieved from the instance,
they were stored into our Bolt DB instance.

The way we chose to store them was to set the issue key (i.e. unique identifier that differentiates a ticket from all others across all 
projects inside a Jira instance) as the key of the pair and the value is the JSON representation of the ticket. This storage method, in conjunction
with the simplicity of Bolt, helped us reduce the size of the database significantly. For small databases, 
using RDBMSs (e.g. Postgres, MySQL, Oracle) is not performance critical, but when the data set is large such as in our study, bottlenecks might start affecting 
the progress of the project. This is mainly due to the structure in which Jira returns the tickets; for example, all issues have comments as 
a field, consisting of an array of structs with body, author, created time etc. Then, each comment has a changelog history attached to it, 
having in turn multiple changelog groups and also each group can have multiple changelog items. Therefore, this long pipeline became a serious 
bottleneck for us once we reached a certain number of tickets because we had relations between all these tables, compared to Bolt which does not 
link any key-value pair to another. Another major point when we decided to proceed with Bolt was that we did not need to query the data at all 
as processing was too complex to convert the logic into an SQL query.

@@

Once the Bolt DB instance is completely populated with all the tickets we were interested in, the analyze command first 
fetches all the issues in the database. The next step is to filter out the tickets based on the following criteria:
\begin{itemize}
  \item tickets are \emph{High Priority} - the ticket status marked as either Blocker, Critical, Major or High;
  \item tickets are closed with a maximum \emph{Time-To-Close} value of 27,000 hours (roughly 3 years) (we added the 
  higher bound as well in order to eliminate outliers retrieved by Jira);
  \item remove outliers for all categories - some fields had extreme values for a small number of tickets (e.g. one ticket
  had over 270,000 comments) introduced by the team behind the project and we needed to exclude them so that the analysis 
  produced correct results (they were removed as they were skewing the results due to erroneous completion of fields).
\end{itemize}

Then, the analysis routine runs in parallel the \emph{seven types of analysis} for answering the research questions 
outlied in Section \ref{intro}, each in its own separate goroutine in order to speed up the process:
\begin{itemize}
  \item \emph{attachments} - checks whether attachments influence \emph{Time-To-Close} and also look at what types 
  of attachments (e.g. screenshots, archive, code snippets) influence \emph{Time-To-Close} the most;
  \item \emph{steps to reproduce} - performs a complex regex to detect steps to reproduce and then checks whether their presence
  influence the \emph{Time-To-Close} of a ticket; the regular expression was adapted after the general guidelines elicited by 
  Bettenburg et al. \cite{bettenburg2008extracting} on how to extract structured data, including enumerations (of which 
  steps to reproduce are a subset);
  \item \emph{stack traces} - runs complex regex for detecting exception stack traces in either summary, description or comments 
  and verifies whether their presence influence Time-To-Close for the ticket; this analysis only inspects projects
  written in Java as it is the most popular proponent built-in of stack traces;
  \item \emph{comments complexity} - loops through all comments and counts all words; then, the tool verifies whether 
  the number of words in comments has an impact on the Time-To-Close for the ticket;
  \item \emph{fields complexity} - same analysis as comments complexity, but only for summary and description wordiness;
  \item \emph{grammar correctness} - it uses Azure Cloud Bing Spell Check API to perform analysis on summary, 
  description and comments; after concatenating everything and making it compatible with the API allowed formats, the tool 
  receives back in the JSON payload not only the number of flagged tokens (i.e. grammar errors), but also their types 
  (e.g. unknown token, misspell);
  \item \emph{sentiment} - uses Google Cloud Platform's Natural Language Processing API to retrieve the 
  sentiment score for summary, description and comments; it first concatenates everything and makes it conform to Google's 
  API and then sends the request, receiving a score from -1 to 1 inclusive (-1 means most negative, 1 means most 
  positive);
\end{itemize}

@@

In this section we present our results to the research questions we set to answer. We have computed charts for all 
independent variables using either scatter plots for continuous variables (e.g. number of comments/\emph{Time-To-Close}) 
or bar charts for categorical variables (e.g. presence of attachments). We also computed statistical analysis for 
all variables to test their significance and rank correlation coefficient and we will present each of them in their 
corresponding subsection.

The answers to our research questions are all statistically significant and they can summarised as follows:
\begin{itemize}
  \item having attachments reduces \emph{Time-To-Close} of a ticket;
  \item having steps to reproduce reduces \emph{Time-To-Close} of a ticket;
  \item having stack traces reduces \emph{Time-To-Close} of a ticket;
  \item more grammar errors imply increase \emph{Time-To-Close} values;
  \item lower sentiment scores result in increased \emph{Time-To-Close} values;
  \item verbose comments result in increased \emph{Time-To-Close} values;
  \item verbose summary and description imply increased \emph{Time-To-Close} values.
\end{itemize}

@@

Using the plot package in our Go tool, we automatically generated the bar chart shown in Figure \ref{attachments}. 
As it can be seen, various types of attachments can produce completely different \emph{times-to-close}. For 
tickets with archive we can see the longest \emph{Time-To-Close} mean while tickets having text attachments are at 
the other end of the spectrum. The smallest mean is for other attachments, which is represented by any file 
that does not have an extension specified in one of the other categories.

We assume that image attachments are screenshots attached to the ticket in order to help developers see and 
subsequently reproduce the bug. Thus, we can clearly see that the difference between the means of \emph{Time-To-Close} 
for tickets without attachments and tickets with screenshots is significantly in favor of the latter. This can 
indicate that developers indeed find screenshots helpful in solving the ticket faster.

Another insightful result is the fact that having text, config and spreadsheet attachments significantly 
reduces the \emph{Time-To-Close} for the ticket. This can imply that developers find value in various steps to reproduce, 
config values used by users when the bug encountered or data in a spreadsheet format (e.g. csv) that can be 
found in such files and helps them lower the \emph{Time-To-Close} for the ticket.

On the other hand, having no attachments does not necessarily mean that the ticket will get closed after 
a longer period of time. As it can be seen in the figure, the mean \emph{Time-To-Close} value for tickets without 
attachments is somewhat neutral in the entire data set.

@@

Thus, our analysis shows that having attachments in most formats, ranging from markdown files to PNG screenshots 
and snippets of code, can help developers solve the tickets faster. We believe that the solution of making attachments
provide the most value to the software project would be if issue tracking systems would make explicit to the reporters 
that uploading a screenshot or a code snippet can help reduce the \emph{Time-To-Close} for the ticket.

@@

As it can be shown in the figure, having steps to reproduce can significantly reduce the \emph{Time-To-Close} 
for the ticket. Even though the data collected does not have such many tickets with steps to reproduce 
(39,988 compared to the total number of 238,286 closed tickets), it is still a valuable finding which shows 
that steps to reproduce are indeed a quality factor for software tickets.

@@

Thus, the answer to our research question is that the presence of steps to reproduce in tickets 
has a positive influence on the \emph{Time-To-Close}. One solution that we propose for maximizing their benefit 
is that issue tracking systems should include hints on the ticket creation form specifying that 
steps to reproduce should be included as developers would be able to fix the bug faster. Moreover, 
issue tracking systems could also automatically add a new system field (like summary or description) 
for any project created and the project maintainers or administrators could make them mandatory for 
when a ticket is created by reporters.

@@

As shown in Figure \ref{stack_traces}, we can see that the presence of exception stack traces can 
reduce the \emph{Time-To-Close} of a ticket in orders of magnitude. This could imply that stack traces help 
developers working on the project localise the code that is running exceptions

@@

A correlational analysis ran on all 201,786 tickets, however, releaved a weak positive relationship with $r = 0.271$ and $p < 0.01$. 
This implies a statistically significant finding - as the number of comments increases, the \emph{Time-To-Close} value of the ticket also increases.

@@

Through calling Microsoft Azure's Bing Spell Check API, one of the best grammar analysis APIs in the industry, we 
managed to calculate the total number of grammar errors in summary, description and comments for a total number 
of 133,689 tickets. As shown in a number of studies, the number of spelling errors can influence the quality of a ticket. 
However, in the work conducted by Schugerl et al. \cite{schugerl2008mining}, the authors found that spelling errors 
can be found in both high and low quality tickets, thus should not be considered as a reliable factor for determining
the Time-To-Close for a ticket. In the rest of this section we will demonstrate the contrary - tickets with more 
grammar errors (spelling errors are just a subset of the types of errors we are investigating) imply lower quality
tickets (i.e. increased \emph{Time-To-Close}).

@@

On the other side of the spectrum, having a more positive score leads to reduced \emph{Time-To-Close} values for tickets. Calefato et al. 
\cite{calefato2015mining} conducted a study where they show that a positive sentiment has the biggest impact on 
the quality of an answer on Stack Overflow, one of the most popular websites related to programming in general. This 
implies that when developers and triagers treat their reporters respectfully and offer support whenever possible, 
the users will feel more welcomed and willing to help report other bugs in the future as well. We also have scores of 
over 0.9, meaning that the actors participating in the discussion (or the reported summary/description) might compliment 
each other for doing a great job.

@@

We manager to answer all seven research questions and find factors that can posittively or 
negatively affect the \emph{Time-To-Close} value of a ticket. All of our results are statistically significant and we can summarise 
them as follows:
\begin{itemize}
  \item the presence of steps to reproduce reduces the \emph{Time-To-Close} for a ticket;
  \item the presence of stack traces reduces the \emph{Time-To-Close} for a ticket;
  \item the presence of attachments reduces the \emph{Time-To-Close} for a ticket - more specifically tickets with screenshots, 
  config files or text files help tickets get closed quicker than those without attachments;
  \item higher numbers of grammar errors imply increased \emph{Time-To-Close} values;
  \item higher sentiment scores imply reduced \emph{Time-To-Close} values for a ticket;
  \item higher numbers of words in comments imply increased \emph{Time-To-Close} values for a ticket;
  \item higher number of words in description and summary increase the \emph{Time-To-Close} values for tickets.
\end{itemize}
